\documentclass[../draft.tex]{subfiles}

\begin{document}
    \chapter{Implementation}\label{c:implementation}
    This chapter describes the details of our backward-directed implementation and how we integrated it into \textsc{FlowDroid}.

    \section{Integration}
    First, we needed a backward interprocedural control-flow graph.
    \textsc{FlowDroid} already contained one for the on-demand aliasing, which suited our needs.
    Next, we need to introduce unconditional taints at sinks and check for the matching access paths at sources.
    The methods for retrieving sources and sinks from a Source Sink Manager have different signatures because, in a forward analysis, access paths only have to match at sinks.
    We added the interface \code{IReversibleSourceSinkManager} extending the \code{ISourceSinkManager}.
    It enforces two additional methods:
    \begin{itemize}
        \item \code{SourceInfo getInverseSinkInfo(Stmt sCallSite,}\\
              \phantom{\code{ceInfo getInverseSinkInfo(}}\code{InfoflowManager manager)}
        \item \code{SinkInfo getInverseSourceInfo(Stmt sCallSite,}\\
              \phantom{\code{Info getInverseSourceInfo(}}\code{InfoflowManager manager, AccessPath ap)}
    \end{itemize}
    \code{getInverseSinkInfo()} returns the necessary information for introducing unconditional taints at sinks, while \code{getInverseSourceInfo()} also matches the access paths at sources.
    All source sink managers needed for the data flow analysis now implement the corresponding interface.
    Note that reversible source sink managers currently do not support the one-source-at-a-time mode.

    For the core flow functions, we created two new classes implementing \code{IInfoflowProblem}.
    \code{BackwardsInfoflowProblem} implements the flow functions described in \autoref{s:flowfunctions}.
    We also refer to this as the main analysis or infoflow analysis\footnotemark{}.
    \footnotetext{%
        The term information flow analysis is sometimes used interchangeably with taint analysis.
        Others use it to describe taint analyses which also support implicit flows \cite{Lerch2014}.
    }
    The flow functions for additional language features are sourced out into rules which are informally described in \autoref{s:rules}.
    The second class is \code{BackwardsAliasProblem} which is the on-demand forward alias analysis.
    We describe the on-demand aliasing in greater detail in \autoref{s:aliasing}.

    After the analysis, the path builder constructs a path out of the leaked taint and its predecessors.
    Because the path builder expects a forward-built taint chain, the path ends up being the wrong way round.
    To hide the fact that we internally searched backward, we also created a \code{BackwardsInfoflowResults} extending \code{InfoflowResults}.
    The implementation is quite simple.
    It overwrites only the \code{addResult()} implementations used by the path builder to save the results and swaps the start and end.
    If full path reconstruction is enabled, it also reverses the path in between.

    \section{Flow-Sensitive Alias Analysis}\label{s:aliasing}
    \textsc{FlowDroid} offers multiple aliasing strategies.
    In this work, we focus on flow-sensitive alias analysis.
    The analysis is formulated as another IFDS problem.
    Basically, it is a forward-directed IFDS analysis using flow function with aliasing rules.
    The main analysis invokes the alias analysis on-demand when it discovers an alias.
    The alias analysis runs independent from the main analysis and later injects found aliases into the main analysis.

    Note that pointer analysis itself is a non-distributive problem \cite{Arzt2017PhD, Spaeth2016}.
    Nonetheless, the alias analysis is encoded in IFDS and we deliberately accept the possibly imprecise results due to the overapproximation.
    Not using the intertwined alias analysis is too imprecise and the cases of overapproximation are rare in practice \cite{Arzt2017PhD}.

    \begin{figure}[tbp]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
            \centering
            \begin{adjustbox}{max width=\columnwidth}
                \begin{lstlisting}[gobble=20]
                    void aliasRule1() {
                        A a = b;
                        b.str = source();
                        sink(a.str);
                    }
                \end{lstlisting}
            \end{adjustbox}
            \caption{Example for alias analysis initiated by rule 1}
            \label{lst:aliasex_a}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.45\textwidth}
            \centering
            \begin{adjustbox}{max width=\columnwidth}
                \begin{lstlisting}[gobble=20]
                    void aliasDefaultRule() {
                        A a = b;
                        a.str = source();
                        sink(b.str);
                    }
                \end{lstlisting}
            \end{adjustbox}
            \caption{Example for alias analysis initiated by the default case}
            \label{lst:aliasex_b}
        \end{subfigure}
        \caption{Normal flow Aliasing examples}
        \label{lst:aliasex}
    \end{figure}

    \paragraph{Handover between the analyses}
    The main analysis discovers aliases at assignments.
    Consider \autoref{lst:aliasex} where two different cases are displayed.
    On the left is a normal flow according to rule 1.
    In line 2, the in taint \code{a.str} produces the outcoming taint \code{b.str}.
    Because the assignment type is a reference type, the backward analysis now recognizes that it possibly missed updates to \code{b.str} below of line 2.
    It invokes the alias analysis with \code{b.str}.
    On the right is a normal flow according to default case.
    In comparison to the left side, this time the assignment in line 2 is swapped.
    At line 2, the main analysis leaves the incoming taint \code{b.str} untouched but notices \code{a} aliases \code{b} below line 2, hence invoking the alias analysis with \code{a.str}.

    The alias analysis searches for missed updates.
    If the analysis found an update, e.g., the taint is on the left side of the assignment, the analysis injects the edge to the statement with the taint into the main analysis' worklist.
    Consider again \autoref{lst:aliasex_a}.
    In line 3, the alias analysis encounters the tainted \code{b.str} on the left side.
    At this point, \code{b.str} gets handed back to the main analysis, following the missed update to find a possible leak. In this case, the leak happens right away.

    \begin{figure}[tbp]
        \centering
        \begin{adjustbox}{max width=\columnwidth}
            \begin{lstlisting}[gobble=16]
                void turnStmtNeeded() {
                    A a = b;
                    String str = b.str;
                    a.str = source();
                    sink(str);
                }
            \end{lstlisting}
        \end{adjustbox}
        \caption{Aliasing example with turn unit}
        \label{lst:turnunit}
    \end{figure}

    \paragraph{Maintaining Flow Sensitivity}
    Arzt solved this in the existing forward implementation using an activation unit.
    This statement marks the update at which the alias gets tainted and can leak at sinks \cite{Arzt2017PhD}.
    This concept does not work for our backward implementation as our alias analysis traverses forward where the taint is already valid.
    Thus we introduce the turn unit.
    The turn unit holds the last observed non-aliasing assignment in the backward analysis.
    When a taint reaches its turn unit in the aliasing analysis, the analysis kills the taint.
    Consider \autoref{lst:turnunit}.
    The newly introduced taint \code{str} at the sink in line 5 also initially has line 5 as a turn unit.
    In line 3, a non-aliasing assignment happens, hence the turn unit is updated.
    In line 2, \code{b.str} is tainted and the alias analysis is invoked with \code{a.str}.
    Without the turn unit, the taint would pass line 3.
    Further in line 4, the taint is handed to the main analysis.
    The main analysis then reports a leak.
    With the turn unit in place, the alias analysis kills the taint in line 4, preventing the false positive.

    % \begin{figure}[tbp]
    %     \centering
    %     \begin{adjustbox}{max width=\columnwidth}
    %         \begin{lstlisting}[gobble=16]
    %             void foo(MyObject pObj) {
    %                 MyObject o = pObj;
    %                 assign(o1, o2);
    %                 assign(o1, o2);
    %                 sink(str);
    %             }

    %             String assign(MyObject o1, MyObject o2) {
    %                 o1.str = o2.str;
    %             }
    %         \end{lstlisting}
    %     \end{adjustbox}
    %     \caption{Summaries And Turn Units}
    %     \label{lst:turnunitsum}
    % \end{figure}
    % \todo{Better example}

    The turn unit is a new field in the \code{Abstraction} class, which is representing a taint.
    A possible drawback of the backward analysis could be the reusability of the IFDS summaries.
    Because the turn unit is part of the taint, IFDS treats equal taints with different turn units as if they have a different context.
    % Consider \autoref{lst:turnunitsum} within the alias analysis.
    % Let us assume IFDS already returned from the call \code{bar(o1);} and created the summary for this context.
    % Later, the analysis observes the same access path with the same context but a different turn unit.
    % Now, IDFS can not apply the summary from \code{bar(o1);}.
    % If turn unit of \code{o1} is inside the callee or one of the transitive callees but the turn unit of \code{o2} is not, we would effectively lose the flow sensitivity as the taint is wrongfully killed.
    % In the reverse case, we would also lose the flow sensitivity because, this time, the turn unit is ignored.

    \section{Rules}\label{s:rules}
    Flow functions can get quite large, complicated to understand and hard to maintain \cite{Lerch2015}.
    To counteract this, \textsc{FlowDroid} outsources certain features into rules.
    These rules also implement the four flow functions and are applied in the main analysis's corresponding flow function.
    In this section, we describe our implementation and informally state the rule behavior.

    \subsection{Source \& Sink Propagation Rule}\label{s:sourcerule}
    In backward analysis, sources act like sinks and vice versa.
    Thus, the Source Propagation Rule records taints flowing into sources and the Sink Propagation Rule unconditionally introduces taints at sinks requiring an \code{IReversibleSourceSinkManager}.

    Notably, the default Source Sink Manager assumes the return value to be tainted.
    Only if the return value is ignored or the method has no return value, the base object is assumed to be tainted.
    While at sinks the base object and parameters are leaked by default \cite{Arzt2017PhD}.
    Thus, starting at sinks might result in more taints per statement.
    We did not notice an impact on the performance in the development phase but the test cases are fairly simple, so the base object creation is often right before the sink call.
    Also, we do not think this impacts the real-world performance of the backward analysis because Arzt evaluation of the forward implementation showed no correlation between runtime and source count \cite{Arzt2017PhD}.
    Whether this is true is sorted out in the evaluation in \autoref{s:realworld}.
    Otherwise, there are more precise Source Sink Managers like the \code{AccessPathBasedSourceSinkManager} where such a problem is not present at all.

    \subsection{Backwards Array Propagation Rule}
    In \textsc{FlowDroid}, array taints are overapproximated by only distinguishing contents and length but not elements. Meaning if one element of an array is tainted, \textsc{FlowDroid} considers all elements tainted.
    Indices are often computed at runtime and thus not available for a static analysis without applying another analysis beforehand. Hence, the approximation is not as severe because we could only track constant indices without large overhead regardless.
    Furthermore, distinguishing elements would increase the domain even more, subsequently increasing the runtime \cite{Arzt2017PhD}.

    The Array Propagation Rule handles \code{ArrayNewExpr}, \code{LengthExpr} and \code{ArrayRef} on the right-hand side.
    \begin{itemize}
        \item \textbf{Array Rule 1}: If the left side's length is tainted and the right side is an \code{ArrayNewExpr}, the outcoming taint is the size local of the \code{ArrayNewExpr}.
        \item  \textbf{Array Rule 2}: If the left side is tainted and the right side is a \code{LengthExpr}, the outcoming taint is the operand of the \code{LengthExpr} with only its length tainted.
        \item \textbf{Array Rule 3:} If the left side is tainted and the right side is an \code{ArrayRef}, the outcoming taint is the array base with only its content tainted.
    \end{itemize}
    The overapproximation of arrays also implies that array taints can not be killed if the left side is an \code{ArrayRef}.

    \subsection{Backwards Exception Propagation Rule}
    The backwards analysis first finds a catch statement \code{\$someVar := @caughtexception}.
    If the left side matches, it sets an exception flag at the taint and propagates it onwards.
    The subsequent propagation then finds the corresponding throw statement.
    \begin{itemize}
        \item \textbf{Exception Rule 1}: On a caught exception expression, derive a new taint with an exception flag set.
        \item \textbf{Exception Rule 2}: If a taint with the exception flag set occurs at a \code{ThrowStmt}, derive taint the operand of the \code{ThrowStmt}.
    \end{itemize}
    The second rule is present in Call and Normal Flow because the throw statement can be inside the same method or in a callee.

    \subsection{Backwards Wrapper Propagation Rule}
    The implementation of this rule is similar to the existing implementation.
    A tainted returned value also needs to be passed into the taint wrapper because of the backward direction.
    The rule calls \code{getInverseTaints()} and thus requires the taint wrapper to implement the \code{IReversibleTaintWrapper} interface.

    \subsection{Backwards Implicit Propagation Rule}
    Flows which are influenced by a condition are called implicit flows.
    A common example is a password check.
    Such a method could return a boolean signifying the password is correct or not.
    Without implicit flows, the taint analysis would be unable to find the path between the password input and the output action.
    The semantics of implicit flows in \textsc{FlowDroid} are that every update flowing into a sink or a sink call inside a conditional branch, even in transitive callees, is considered a leak.

    The existing forward-directed implementation derives a wildcard taint\footnotemark{} and propagates it until the conditional branch is left again at the immediate postdominator.
    \footnotetext{A wildcard taint is an taint without an object or access path but a wildcard appended.}
    This behavior does not scale well because the semantics demand tainting every update and following every call in conditionals leading to many unneccessary taints never reaching a sink.
    The bad scalability is not an issue specific to \textsc{FlowDroid} but is present for most analyses providing implicit flows \cite{King2008}.

    We also use the wildcard taint but with a different semantic.
    Backward, the wildcard signifies a conditional update and is propagated further to find the corresponding condition.
    Thus, the branching factor inside a conditional branch should be lower.
    To detect a conditonal, we push and pop the dominator on the taints.
    Whenever an update is conducted to a taint carrying a dominator, the wildcard taint is derived.
    However, it is not that easy to reconstruct the conditional branches while traversing the exploded supergraph.
    Consider the code in \autoref{lst:impflow}.
    We expect from the analysis to find the path from line 2 to line 10.
    First of all, starting at the sink it is unknown whether a conditional influenced the call to the \code{transitiveSink()} method.
    Also, assuming we created a taint representing the sink call, consider the return edge into \code{foo()}.
    The edge is already inside a conditional branch, thus the dominator is per-definition line 4.
    We cannot use the dominator as a indication whether a taint enters a conditional in such cases.
    Both of those challenges are not a data-flow problem but rather a control-flow graph reachability problem.
    We extended the backward interprocedural control-flow graph with two methods:
    \begin{itemize}
        \item \code{List<Unit> getConditionalBranchesInterprocedural(Unit unit)}
        \item \code{List<Unit> getConditionalBranchesIntraprocedural(Unit unit)}
    \end{itemize}
    The first one is used for sink calls.
    It traverses the interprocedural graph using a worklist algorithm to find all possibly reachable conditional statements.
    We then use those found conditionals to derive sink taints with the correct conditional dominator.
    The second one only returns the conditional statements in the above the unit but does not follow transitive call sites.
    For example, the second method is used to reconstruct 

    \begin{figure}[tbp]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
            \centering
            \begin{lstlisting}[gobble=16]
                void foo() {
                    int tainted = source();
                    if (tainted == 42) {
                        int x = 0;
                        transitiveSink();
                    }
                }

                void transitiveSink() {
                    sink();
                }
            \end{lstlisting}
        \end{subfigure}
        \caption{Implicit Flow Challenges}
        \label{lst:impflow}
    \end{figure}

    \subsection{Backwards Strong Update Rule}
    Until now, we always assumed that a taint is only affected if the variable occurs in a statement.
    However, with aliasing, this gets more complicated.
    A taint could not match the left side and, thus, is propagated over the statement according to the default rule of normal flow, but the taint is an alias of the left side and should have been killed.
    Also, we can not just link aliases to taints for such strong updates because that would violate the flow functions' distributiveness property.

    In this case, \textsc{FlowDroid} falls back to Soot's must-aliasing analysis. However, the must-aliasing analysis is only intraprocedural. Thus, strong updates split over methods are not detected and produce a false positive.

    Backward, the first observed update is the correct one. We treat a must-alias like a regular match:
    \begin{itemize}
        \item \textbf{Strong Update Rule}: If the incoming taint must-aliases the left side, then apply the normal flow rules just as if the left side was tainted.
    \end{itemize}

    \subsection{Backwards Clinit Rule}\label{s:clinitrule}
    \code{<clinit>} is a special method in the JVM and stands for class loader init.
    The compiler generates the method and calls it implicitly. Examples of statements that get compiled into clinit are in \autoref{lst:clinit_examples}.
    The invokation is implicit at the class's initialization phase and is executed at most once for each class\footnotemark{}.
    \footnotetext{Source: \url{https://docs.oracle.com/javase/specs/jvms/se8/html/jvms-2.html\#jvms-2.9} (visited on 17.04.2021)}
    SPARK, the default call graph algorithm of \textsc{FlowDroid}, overapproximates the \code{<clinit>} behavior.
    It adds an edge to \code{<clinit>} at each statement containing a \code{StaticFieldRef}, \code{StaticInvokeExpr} or \code{NewExpr}\footnotemark{}.
    \footnote{\url{https://github.com/soot-oss/soot/blob/59931576784b910a7d38f81910b7313aa2feafea/src/main/java/soot/jimple/toolkits/callgraph/OnFlyCallGraphBuilder.java\#L969} (visited on 17.04.2021)}

    \begin{figure}[tbp]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
            \centering
            \begin{lstlisting}[gobble=16]
                class ClinitClass1 {
                    public static string str = source();
                }
            \end{lstlisting}
            \caption{static variable initialization}
            \label{lst:clinit_examples_a}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.45\textwidth}
            \centering
            \begin{lstlisting}[gobble=16]
                class ClinitClass2 {
                    static {
                        ClinitClass2.sink();
                    }
                }
            \end{lstlisting}
            \caption{static block}
            \label{lst:clinit_examples_b}
        \end{subfigure}
        \caption{Examples of statements being in \code{<clinit>}}
        \label{lst:clinit_examples}
    \end{figure}

    The need for this rule is rooted in the IFDS solver of \textsc{FlowDroid}.
    The solver decides whether to use Normal Flow or Call Flow by calling \code{isCallStmt(Unit u)} on the interprocedural control-flow graph generated by Soot.
    Internally, this method calls \code{containsInvokeExpr()} on the \code{Unit} object. \code{containsInvokeExpr()} for \code{AssignStmt} only returns true if the right-hand side is an instance of \code{InvokeExpr}.
    Consequently, the calls to \code{<clinit>} from \code{AssignStmt}s with \code{NewExpr} or \code{StaticFieldRef} on the right side are missed.

    The Backwards Clinit Rule manually injects an edge to the \code{<clinit>} method in the infoflow solver when appropriate during the analysis.
    Also, it lessens the overapproximation of SPARK by carefully choosing whether to inject the edge. The rule works as follows:
    \begin{itemize}
        \item \textbf{Clinit Rule 1}: If the tainted static variable is a field of the class where the current statement is located in, do not inject because we will at least encounter a \code{NewExpr} of the same class further in the call graph.
        \item \textbf{Clinit Rule 2}: Else if the tainted static variable matches the \code{StaticFieldRef} on the right hand side: Inject the edge because we can not be sure whether we see another edge to \code{<clinit>}.
        \item \textbf{Clinit Rule 3}: Else if the class of the tainted static variable matches the class of the \code{NewExpr}: Inject the edge because we can not be sure whether we see another edge to \code{<clinit>}.
    \end{itemize}
    The behavior is still an overapproximation, of course.
    A more precise solution would require bookkeeping of every class's last observation equal to the first occurrence in the code.

    In the existing implementation, there is no such explicit modelling.
    As taints are introduced at sources, if the source statement is inside a static initialization like in \autoref{lst:clinit_examples_a}, the propagation starts inside the \code{<clinit>} method.
    Then unbalanced problem handling of the solver propagates the taint into the callers of \code{ClinitClass1 <clinit>}.
    This behavior allows the forward analysis to detect leaks originating from static variable initializations.
    Leaks from inside a static blocks as shown in \autoref{lst:clinit_examples_b} can be missed by the forward analysis because it might no call statement with an edge to \code{<clinit>}.
    The Clinit Rule could also be ported to the forward analysis but with a larger overapproximation. Unlike backward, there is no guarantee that there will be another edge to \code{<clinit>} if the statement is in the same class as the \code{<clinit>} method. Thus, it would inherit the entire overapproximation from the SPARK call-graph algorithm.


    \subsection{Other Rules}
    Skip System Class Rule and Stop After First K Flows Rule are not direction-dependent.
    Both are shared with the forwards search and therefore use the existing implementation in \textsc{FlowDroid}.

    % Typing Propagation Rule has no backwards equivalent. We decided to implement type checking in the infoflow problem instead.

    \section{Other Components}
    \subsection{Taint Wrappers}\label{s:taintwrapper}
    \textsc{FlowDroid} already has an interface \code{IReversibleTaintWrapper} for taint wrappers providing inversed summaries.
    The \code{SummaryTaintWrapper} using StubDroid's summaries already implemented this interface.
    For the \code{EasyTaintWrapper}, we contributed the inverse implementation.
    Its implementation follows simple rules which cover most cases\cite{Arzt2017PhD}.
    The rules are inverted to:
    \begin{itemize}
        \item If the return value is tainted, taint the object and the parameters.
        \item If the base object is tainted, taint all parameters.
    \end{itemize}

    \begin{figure}[tbp]
        \centering
        \begin{adjustbox}{max width=\columnwidth}
            \begin{lstlisting}[gobble=16]
                char[] tainted = source();
                StringBuilder sb = new StringBuilder();
                sb.append(tainted, offset, len);
                sb.append("untainted");
                sink(sb.toString());
            \end{lstlisting}
        \end{adjustbox}
        \caption{Easy Taint Wrapper Example}
        \label{lst:etw}
    \end{figure}

    Note that these simple rules are disadvantageous for the backward direction the more parameters a method has.
    Consider the code snippet in \autoref{lst:etw}, especially line 3.
    Forwards, \code{tainted} is the incoming taint and the \code{EasyTaintWrapper} produces the taint set \code{\{tainted, sb\}}.
    Backward, the incoming taint is \code{sb} and the taint wrapper produces four taints \code{\{sb, tainted, offset, len\}}.
    Luckily, most methods supported by the \code{EasyTaintWrapper} have less than three arguments.
    Alternatively, using the more precise \code{SummaryTaintWrapper} eliminates this problem. 
    We compare both taint wrappers in \autoref{s:droidbenchevaluation}.

    \subsection{Native Call Handler}
    The native call handler of \textsc{FlowDroid} handles two methods: \code{System#arraycopy} and \code{reflect.Array#newArray}.
    The handling of \code{System#arraycopy} is direction-dependent.
    Thus, we adapted the existing implementation and reversed the logic of \code{System#arraycopy} to reflect the analysis direction.

    \subsection{Code Optimizer: AddNOPStmts}
    Before starting the analysis, \textsc{FlowDroid} applies code optimization to the interprocedural call graph.
    By default, dead code elimination and within constant value propagation is performed.
    Those are also applied before backward analysis,but we needed another code optimizer to handle an edge case in backward analysis.

    First, consider the \code{static2Test} test case in the \code{StatictTestCode} class of \textsc{FlowDroid} in \autoref{lst:static2TestJava}.
    The method is also the entry point for the analysis, is static and does not have any parameters.
    The same is true for the source \code{TelephonyManager#getDeviceId}.
    Due to the first condition, \code{static2Test} has no identity statements and because of the second condition, there are also no assign statements before the source statement in Jimple.
    Therefore the source statement is the first statement in the graph.
    Next, a detail of \textsc{FlowDroid}'s IFDS solver is important.
    The Return and CallToReturn flow function is only applied if a return site is available.
    When traversing backward, the source statement is the last and thus has no return sites.
    Now, the taints flowing into source methods are registered in the Call To Return flow function.
    Altogether, leaks are missed if the source statement is the first statement.

    Moving the detection of incoming taints flows into sources from the CallToReturn to the Call flow function was not an option because by default source methods are not visited and changing this would require multiple changes in the existing implementation.
    Our solution is to add a NOP statement in such cases before the analysis.
    Due to the entry points being known beforehand, the overhead is negligible.

    \begin{figure}[tbp]
        \centering
        \begin{subfigure}[b]{\textwidth}
            \begin{lstlisting}[gobble=16]
                public static void static2Test() {
                    String tainted = TelephonyManager.getDeviceId();
                    ClassWithStatic static1 = new ClassWithStatic();
                    static1.setTitle(tainted);
                    ClassWithStatic static2 = new ClassWithStatic();
                    String alsoTainted = static2.getTitle();

                    ConnectionManager cm = new ConnectionManager();
                    cm.publish(alsoTainted);
                }
            \end{lstlisting}
            \caption{Java}
        \end{subfigure}
        \qquad
        \begin{subfigure}[b]{\textwidth}
            \begin{lstlisting}[language=Jimple, gobble=16]
                public static void static2Test() {
                    tainted = staticinvoke <soot.jimple.infoflow.test.android.TelephonyManager: java.lang.String getDeviceId()>(); // Line 2 in (a)

                    // [...]

                    virtualinvoke cm.<soot.jimple.infoflow.test.android.ConnectionManager: void publish(java.lang.String)>(alsoTainted); // Line 9 in (a)

                    return;
                }
            \end{lstlisting}
            \caption{Jimple}
        \end{subfigure}
        \caption{static2Test Code}
        \label{lst:static2TestJava}
    \end{figure}

    \subsection{Omit Aliasing Optimization}
    \begin{figure}[tbp]
        \centering
        \begin{adjustbox}{max width=\columnwidth}   
            \begin{lstlisting}[gobble=16]
                public void JSONTest() {
                    String tainted = TelephonyManager.getDeviceId();
                    JSONObject jsonO = new JSONObject();
                    jsonO.put("str", tainted);
                    ConnectionManager cm = new ConnectionManager();
                    cm.publish(jsonO.toString());
                }
            \end{lstlisting}
        \end{adjustbox}
        \caption{JSON Test Code}
        \label{lst:json}
    \end{figure}

    Recall \autoref{s:complexity} where we explained our backward implementation benefits from a right-to-left order.
    The \code{StringBuilder} however can alias and the alias search offsets the advantage.
    We use the observation that Jimple does not reuse the local variables when compiled from Java bytecode and all locals pointing to the same \code{StringBuilder} instance except one are not reused either.
    Thus, we apply a live variables analysis for the base object between the current statement and the turn unit as a preanalysis.
    Only if the preanalysis finds an use, the alias analysis is used.
    The preanalysis needs around 100$\mu s$ runtime per query.
    Based on our observations, it should rarely find an use.
    We reviewed the optimization using the well-known \code{org.json}\footnotemark{} library.
    \footnotetext{https://mvnrepository.com/artifact/org.json/json/20210307 (visited 17.04.2021)}
    The library makes use of \code{StringWriter}, which internally uses \code{StringBuffer}. A \code{StringBuffer} has the same chaining in place for the \code{append()} and \code{insert()} methods.
    For the test case shown in \autoref{lst:json}, we observed 58 misses and 0 hits while spending 0.0337 seconds in the preanalysis until the analysis found the first flow.
\end{document}